---
title: "Employment Study(Case Study 2)"
author: "Joe Schueder/Thejas Prasad"
date: "November 24, 2017"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(maps)
library(mapproj)
library(rvest)
library(plyr)
library(reshape2)
library(tidyr)
library(ggplot2)
library(ggthemes)
```

#####Prepared by Joe Schueder and Thejas Prasar for "WAM (We are Motivated), Inc.

WAM Inc. would like an analysis of where to place their next set of employment training centers to market job productivity training.  The primary criteria that WAM uses when creating new centers is motivation levels/procrastination levels, income averages, and education levels.  They are also interested in how well developed the country of a potential center is.

The primary goal of this analysis was to understand the current climate for employment training centers around the world. This was done by analyzing employee data from several countries which included relevant procrastination scores, education, occupation, and income levels.  This data was combined with the internationally recognized HDI index.

The analysis is shown below. 

Data Preparation to read in and clean data was done.

```{r dataandclean}
#1a
getwd()
setwd("C:/Users/jjschued/Documents")
procrast <- read.csv(file = "Procrastination.csv", sep=",", header=TRUE)

#1a
# number of rows and columns
dim(procrast)
#1b column names

#code below is for viewing all column names befoe re-lableing
#colnames(procrast)

#relabling column names for readability
colnames(procrast) <- c("Age","Gender","Kids","Edu","WorkStatus","AnnualIncome","CurrentOccup","Timeinposyr","Timeinposmn","Communitysize","Countryres","MaritalStatus","Numsons","Numdaughters","DP1","DP2","DP3","DP4","DP5","AIP1","AIP2","AIP3","AIP4","AIP5","AIP6","AIP7","AIP8","AIP9","AIP10","AIP11","AIP12","AIP13","AIP14","AIP15","GP1","GP2","GP3","GP4","GP5","GP6","GP7","GP8","GP9","GP10","GP11","GP12","GP13","GP14","GP15","GP16","GP17","GP18","GP19","GP20","SWLS1","SWLS2","SWLS3","SWLS4","SWLS5","UCONSPRO","OTHCONSPRO")

#2c i get time in service out of scientific notation
procrast$Timeinposyr <- as.integer(format(procrast$Timeinposyr, scientific = F))
# Turning 999 into NA
procrast[,"Timeinposyr"][procrast[,"Timeinposyr"]==999] <- NA
#2 c ii converting Male and Female into numeric representation
procrast$Numsons <- as.character(procrast$Numsons)

procrast$Numsons[procrast$Numsons == "Male"] <- "1"
procrast$Numsons[procrast$Numsons == "Female"] <- "2"
#2ciii
#get rid of 0 and blank and convert them to NA
procrast$Countryres <- as.character(procrast$Countryres)
procrast$Countryres[procrast$Countryres == "0"] <- "NA"
procrast$Countryres[procrast$Countryres == " "] <- "NA"

#2civ
#fix the occupation column NAs blanks, group
procrast$CurrentOccup <- as.character(procrast$CurrentOccup)
procrast$CurrentOccup[procrast$CurrentOccup == "0"] <- "NA"
procrast$CurrentOccup[procrast$CurrentOccup == "please specify"] <- "NA"
procrast$CurrentOccup = toupper(procrast$CurrentOccup)
#grouping like together now
procrast$CurrentOccup <- gsub("TEACHER.*","TEACHER", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("TECH.*","TECHNICAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("WRIT*","WRITER", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("STUDENT.*","STUDENT", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("*.STUDENT","STUDENT", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("SYSTEM.*","TECHNICAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("SOFTWARE.*","TECHNICAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("SELF.*","SELF EMPLOYED", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("SALES.*","SALES", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("RESEARCH.*","RESEARCH", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("REAL ESTATE.*","REAL ESTATE", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("REALTOR.*","REAL ESTATE", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("PROGRAMMER.*","TECHNICAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("IT .*","TECHNICAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("COMPUTER.*","TECHNICAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("DIRECTOR.*","DIRECTOR", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("FINANC.*","FINANCIAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("INSURANCE.*","INSURANCE", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("INFORMATION.*","TECHNICAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("MEDICAL.*","MEDICAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("LEGAL.*","LEGAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("LAW.*","LEGAL", procrast$CurrentOccup)
procrast$CurrentOccup <- gsub("OFFICE.*","OFFICE", procrast$CurrentOccup)
#code below can be used to view all occupations
# kable(unique(procrast$CurrentOccup), "html", padding = 2, longtable = TRUE) %>%
# kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

#2d CHECK FIELD TYPES nothing wrong for now will convert later if needed
#code commendted below can be uncommented to view data typess
#str(procrast)

#2e Get the means of the procrastination indexes

procrast$dpmean <- (procrast$DP1 + procrast$DP2 + procrast$DP3 + procrast$DP4 + procrast$DP5)/5

procrast$aipmean <- (procrast$AIP1 + procrast$AIP2 +procrast$AIP3 +procrast$AIP4 +procrast$AIP5 +procrast$AIP6 +procrast$AIP7 +procrast$AIP8 +procrast$AIP9 +procrast$AIP10 +procrast$AIP11 +procrast$AIP12 +procrast$AIP13 +procrast$AIP14 + procrast$AIP15) /15

procrast$gpmean <- (procrast$GP1 + procrast$GP2 + procrast$GP3 + procrast$GP4 + procrast$GP5 + procrast$GP6 + procrast$GP7 + procrast$GP8 + procrast$GP9 + procrast$GP10 + procrast$GP11 + procrast$GP12 + procrast$GP13 + procrast$GP14 + procrast$GP15 + procrast$GP16 + procrast$GP17 + procrast$GP18 + procrast$GP19 + procrast$GP20)/20

procrast$swlsmean <- (procrast$SWLS1 + procrast$SWLS2 + procrast$SWLS3 + procrast$SWLS4 + procrast$SWLS5)/5

#code below for viewing table contents
kable(head(procrast), "html", padding = 2, longtable = TRUE) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

## 3a Getting data about human development index from wikipedia

URL <- 'https://en.wikipedia.org/wiki/List_of_countries_by_Human_Development_Index#Complete_list_of_countries'

#very high index countries
  HDIURL <- URL %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/div/div[5]/table') %>%
  html_table(fill = TRUE)
  HDITABVH <- HDIURL[[1]]

#labeling useful columns  
colnames(HDITABVH)[1:5] <- c("Rank", "RankChng", "Country", "HDI", "HDICHNG")

#limiting columns and categorizing this table
HDITABVHS <- HDITABVH[1:5]

HDITABVHS$HDICAT <- "Very High"

#getting rid of extra rows
HDITABVHS <- HDITABVHS[-c(1, 2, 3, 30, 31), ]

#high index countries
  HDIURL <- URL %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/div/div[6]/table') %>%
  html_table(fill = TRUE)
  HDITABH <- HDIURL[[1]]

#change column names
  
colnames(HDITABH)[1:5] <- c("Rank", "RankChng", "Country", "HDI", "HDICHNG")
HDITABHS <- HDITABH[1:5]
HDITABHS$HDICAT <- "High"
  
#getting rid of extra rows including one labled world which is not a country
HDITABHS <- HDITABHS[-c(1, 2, 3, 32, 33, 56), ]

#medium index countries
  HDIURL <- URL %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/div/div[7]/table') %>%
  html_table(fill = TRUE)
  HDITABM <- HDIURL[[1]]
colnames(HDITABM)[1:5] <- c("Rank", "RankChng", "Country", "HDI", "HDICHNG")
  
HDITABMS <- HDITABM[1:5]
HDITABMS$HDICAT <- "Medium"

#getting rid of extra rows
HDITABMS <- HDITABMS[-c(1, 2, 3, 24, 25), ]

#low index countries
  HDIURL <- URL %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/div/div[8]/table') %>%
  html_table(fill=TRUE)
  HDITABL <- HDIURL[[1]]
  
  colnames(HDITABL)[1:5] <- c("Rank", "RankChng", "Country", "HDI", "HDICHNG")
  HDITABLS <- HDITABL[1:5]
  HDITABLS$HDICAT <- "Low"
  
#getting rid of extra rows
HDITABLS <- HDITABLS[-c(1, 2, 3, 25, 26), ]
  
# The commented code below was used to explore the data to eliminate bad rows
# kable(HDITABVHS, "html", padding = 2, longtable = TRUE, row.names = TRUE) %>%
# kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
# 
# kable(HDITABHS, "html", padding = 2, longtable = TRUE, row.names = TRUE) %>%
# kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
# 
# kable(HDITABMS, "html", padding = 2, longtable = TRUE, row.names = TRUE) %>%
# kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
# 
# kable(HDITABLS, "html", padding = 2, longtable = TRUE, row.names = TRUE) %>%
# kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
 
#3b merge all the tables together
hdicomb <- rbind(HDITABVHS, HDITABHS, HDITABMS, HDITABLS)

kable(hdicomb, "html", padding = 2, longtable = TRUE, row.names = TRUE) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

dim(hdicomb)

#3c
#write table to csv(6a)
#setwd("C:/Users/jjschued/Documents")
write.table(hdicomb, file = "hdicomb.csv", col.names = TRUE, row.names = FALSE, sep =",")

#merge procrastination and and hdicomb
colnames(hdicomb)[3] <- "Countryres"
procrastcomb <- merge(procrast, hdicomb, by=c("Countryres"), all = FALSE )



kable(head(procrastcomb), "html", padding = 2, longtable = TRUE, row.names = TRUE) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

#4a remove everone under 18 and over 65 since this is for a work analysis since people typically stop working at 65

procrastcomb <- procrastcomb[procrastcomb$Age > 18, ]
procrastcomb <- procrastcomb[procrastcomb$Age < 65, ]

#convert HDI to numeric

procrastcomb$HDI <- as.numeric(format(procrastcomb$HDI, scientific = F))

#write table to csv(6b)
write.table(procrastcomb, file = "procrastcomb.csv", col.names = TRUE, row.names = FALSE, sep =",")

```



Summary statistics of the important parameters are presented below for all of the employee survey responses.  This gives a baseline for what an employer would expect on average for an employee across the world.  Following the summary statistics there are several visuals that help represent this information.  It shows that motivation or procrastination levels are an evenly distributed value with most people falling in the middle.  Income, on the other hand is left skewed with a few clusters of higher income employees, but many others on the lower end of the pay spectrum.  Following the visualizations, some additional summary information is presented about the respondents.  These include: gender, employment status,occupation titles, country of residence, and procrastination perception.  The gender count is evenly split, while most workers are full time.  Occupation title was a very inconsistent field therefore there is a concentration of frequent titles at the top and quickly trailing off to many additional titles with only one respondent. This is sorted in descending order to easily identify top titles.  Next, Country of residence is displayed in descending order as well and it can be noted many respondents are from the United States. Finally, an analysis of perceptions are presented.  Shown are comparisons of where perceptions match and where they do not match between the respondents and others. This shows that most people consider themselves procrastinators.

```{r prelimanalysi}
# 4b summary statistics of procrastcomb
sumtable <- summary(procrastcomb$Age)
Statistics <- c("Min.", "1st Qu.",  "Median", "Mean", "3rd Qu.", "Max.", "NA's" )
Age <- c(sumtable[1], sumtable[2],  sumtable[3], sumtable[4], sumtable[5], sumtable[6], sumtable[7] )

sumtable <- summary(procrastcomb$AnnualIncome)
Income <- c(sumtable[1], sumtable[2],  sumtable[3], sumtable[4], sumtable[5], sumtable[6], sumtable[7] )

sumtable <- summary(procrastcomb$HDI)
HDI <- c(sumtable[1], sumtable[2],  sumtable[3], sumtable[4], sumtable[5], sumtable[6], sumtable[7] )
sumtable <- summary(procrastcomb$aipmean)
AIPmean <- c(sumtable[1], sumtable[2],  sumtable[3], sumtable[4], sumtable[5], sumtable[6], sumtable[7] )
sumtable <- summary(procrastcomb$dpmean)
DPmean <- c(sumtable[1], sumtable[2],  sumtable[3], sumtable[4], sumtable[5], sumtable[6], sumtable[7] )
sumtable <- summary(procrastcomb$gpmean)
GPmean <- c(sumtable[1], sumtable[2],  sumtable[3], sumtable[4], sumtable[5], sumtable[6], sumtable[7] )
sumtable <- summary(procrastcomb$swlsmean)
SSWLSmean <- c(sumtable[1], sumtable[2],  sumtable[3], sumtable[4], sumtable[5], sumtable[6], sumtable[7] )
sumdf <- data.frame(Statistics, Age, Income, HDI, AIPmean, DPmean, GPmean, SSWLSmean)
kable(sumdf, "html", padding = 2, longtable = TRUE, row.names = FALSE, digits = 2) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

#Histograms of two vairables above.  AIP is very much a bell shape curve and evenly distributed, while income is a left skewed curve with most falling on the lower end of the income scale
hist(procrastcomb$aipmean, col="Green", ylab ="AIP Mean", xlab
="FREQUENCY", main = "AIP Mean Histogram")
hist(procrastcomb$AnnualIncome, col="Blue", ylab ="Annual Income", xlab
="FREQUENCY", main = "Annual Income Histogram")

## 4c frequencies of Gender, work Status, Occupation
Gendersum <- count(procrastcomb, 'Gender')
WStatussm <- count(procrastcomb, 'WorkStatus')
OccuStSm <- count(procrastcomb, 'CurrentOccup')

colnames(Gendersum)[2] <- "Frequency"
colnames(WStatussm)[2] <- "Frequency"
colnames(OccuStSm)[2] <- "Frequency"

OccuStSm <- OccuStSm[order(-OccuStSm$Frequency),]

kable(Gendersum, "html", padding = 2, longtable = TRUE, row.names = FALSE) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

kable(WStatussm, "html", padding = 2, longtable = TRUE, row.names = FALSE) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

kable(head(OccuStSm, 20), "html", padding = 2, longtable = TRUE, row.names = FALSE) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

# 4d Count of respondents by country
Countrycnt <- count(procrastcomb, 'Countryres')
colnames(Countrycnt)[2] <- "Frequency"
Countrycnt <- Countrycnt[order(-Countrycnt$Frequency),]
kable(Countrycnt, "html", padding = 2, longtable = TRUE, row.names = FALSE) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

#str(procrastcomb)
#4e matched perceptions
procrastcomb$UCONSPRO <- as.character(procrastcomb$UCONSPRO )
procrastcomb$OTHCONSPRO <- as.character(procrastcomb$OTHCONSPRO)
#where repsonses match
procrastmatch <- procrastcomb[which(procrastcomb$UCONSPRO == procrastcomb$OTHCONSPRO), ]
promatchsum <- count(procrastmatch, 'UCONSPRO')
colnames(promatchsum)[1] <- "Responses"
colnames(promatchsum)[2] <- "#Matches"
kable(promatchsum, "html", padding = 2, longtable = TRUE, row.names = FALSE, caption ="Personal perception procrastination responses that match others perceptions") %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
#where responses do not match
procrastmismatch <- procrastcomb[which(procrastcomb$UCONSPRO != procrastcomb$OTHCONSPRO), ]
promismatchsum <- count(procrastmismatch, 'UCONSPRO')
colnames(promismatchsum)[1] <- "Responses"
colnames(promismatchsum)[2] <- "#MisMatches"
kable(promismatchsum, "html", padding = 2, longtable = TRUE, row.names = FALSE, caption ="Personal perception procrastination responses that do not match others perceptions") %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

To get to where the greatest need is for the productivity training analysis must be shown at the country level.  When top most procrastinating countries compared by two different procrastination survey indices there emerges that there are eight countries that are in both indices. These are the Dominican Republic, Qatar, Panama, Ecuador, Sri Lanka, Turkey, Uruguay, and Iran. Many of these countries are in the HDI Category high. This seems like they are countries that would benefit most from the centers.  If, the company wished to focus more on countries with higher disposable incomes it might be best to limit the data to HDI category Very High.  If this restriction is made the top countries would be Qatar, Austria, Poland, Germany, Hong Kong, and Iceland.

```{r detaianalysis}
#5b barchart of top 15 dp

top15dp <- aggregate(procrastcomb[, 62], by = list(procrastcomb$Countryres, procrastcomb$HDICAT), mean, na.rm=TRUE)
colnames(top15dp) <- c("Country", "HDICat", "DPmean")

#get the dp countries for very high hdi
top15dpvh <- subset(top15dp, HDICat=='Very High' )
top15dpvh <- top15dpvh[order(-top15dpvh$DPmean),]
#top15dpvh

top15dp <- top15dp[order(-top15dp$DPmean),]
#top15dp
top15dp <- top15dp[1:15,]
#top15dp

#write table to csv(6c)
#setwd("C:/Users/jjschued/Documents")
write.table(top15dp, file = "top15dp.csv", col.names = TRUE, row.names = FALSE, sep =",")

 ggplot(top15dp, aes(reorder(Country, DPmean),DPmean))+
 geom_bar(aes(fill=HDICat), stat= "identity", width = .4, position = position_dodge(width = .9))+
 ggtitle("Top 15 DP Mean Countries")+
 theme(plot.title = element_text(hjust = 0.5), legend.position="none",)+
 xlab("Country")+
 ylab("DP Means")+
 theme(axis.text.y=element_text(hjust = .5, vjust = .5))+
  theme_solarized()+
 coord_flip()

#5c barchart of top 15 aip
top15aip <- aggregate(procrastcomb[, 63], by = list(procrastcomb$Countryres, procrastcomb$HDICAT), mean, na.rm=TRUE)
colnames(top15aip) <- c("Country", "HDICAT","AIPmean")
top15aip <- top15aip[order(-top15aip$AIPmean),]
# top15aip
#write table to csv(6c)
#setwd("C:/Users/jjschued/Documents")
write.table(top15aip, file = "top15aip.csv", col.names = TRUE, row.names = FALSE, sep =",")

#get the aip countries for very high hdi
top15aipvh <- subset(top15aip, HDICAT=='Very High' )
top15aipvh <-top15aipvh[order(-top15aipvh$AIPmean),]
#top15aipvh

top15aip <- top15aip[1:15,]
#top15aip

 ggplot(top15aip, aes(reorder(Country, AIPmean),AIPmean))+
 geom_bar(aes(fill=HDICAT), stat= "identity", width = .4, position = position_dodge(width = .9))+
 ggtitle("Top 15 AIP Mean Countries")+ 
 theme(plot.title = element_text(hjust = 0.5), legend.position="none",)+
 xlab("Country")+
 ylab("AIP Means")+ 
 theme(axis.text.y=element_text(hjust = .5, vjust = .5))+
 theme_solarized()+
 coord_flip()


```


WAM also requested supplemental analysis regarding how to target advertising for their series of trainings.  Their interest was for ages, income, life satisfaction, and HDI categories of the targeted countries. The best relationship seems to be income to age.  There is more income at older ages. Other relationships were not consequential. Therefore, it is suggested to examine targeting older individuals. 


```{r relationships}
 #5d relationship between age and income
 
 ggplot(procrastcomb, aes(x= procrastcomb$Age, y=procrastcomb$AnnualIncome))+
 geom_point(aes(col = procrastcomb$Gender))+
   guides(col=guide_legend(title="Gender"))+
geom_smooth(method=lm)+
 ggtitle("Age and Income Relationship")+ 
 theme(plot.title = element_text(hjust = 0.5), legend.position="none",)+
 xlab("Age")+
 ylab("Income")+ 
 theme(axis.text.y=element_text(hjust = .5, vjust = .5))+
 theme_solarized()

 
 #5e relationship between Life satisfaction and HDI
 
 ggplot(procrastcomb, aes(x= procrastcomb$HDI, y=procrastcomb$swlsmean))+
 geom_point()+
geom_smooth(method=lm)+
 ggtitle("HDI and Life Satisfaction relationship")+ 
 theme(plot.title = element_text(hjust = 0.5), legend.position="none",)+
 xlab("HDI")+
 ylab("Life Satisfaction Mean")+ 
 theme(axis.text.y=element_text(hjust = .5, vjust = .5))+
 theme_solarized() 
 
 #HDI cat and life satisfaction

 ggplot(procrastcomb, aes(x= procrastcomb$swlsmean, y=procrastcomb$HDICAT))+
 geom_bar(aes(fill=HDICAT), stat= "identity", width = .4, position = position_dodge(width = .9))+
 ggtitle("HDI and Life Satisfaction relationship")+
 theme(plot.title = element_text(hjust = 0.5), legend.position="none",)+
 xlab("Life Satisfaction Mean")+
 ylab("HDICat")+
 theme(axis.text.y=element_text(hjust = .5, vjust = .5))
 theme_solarized()

 
 
```
In conclusion, though beyond the United States and a few other countries sample sizes are small the results suggest that there are markets that warrant additional analysis outside the scope of this project.  Those include: central Europe, Hong Kong, and Qatar.  Further surveys of employment age individuals should be gathered for all HDI Category-Very High to make better decisions about placement of their centers. It is important to place the centers in high-income areas in order to maximize exposure to customers with disposable income.   
